### 2022-06-21 src/ms.tex
Almost done with Chapter 2

% Define document class
\documentclass[12pt]{report}
\usepackage{aas_macros}

@@

% Figures & tables 
\usepackage{graphicx,xcolor}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs} % much better looking tables

@@

@@

@@

% Custom command
\newcommand{\ud}{\,\mathrm{d}}
\renewcommand{\vec}[1]{\boldsymbol{\mathbf{#1}}}

\hypersetup{
    colorlinks=true,
    allcolors=cyan,
}

@@

The assumptions that existing methods for modeling microlensing events rely on
are often opaque and unquestioned. Discussions on model ``degeneracies''
\citep{2014MNRAS.437.4006S,2019AJ....157...23H,2018AcA....68...43S,2009MNRAS.393..816D},
correlated noise \citep{2015ApJ...812..136B,2019MNRAS.488.3308L} and model
comparison \citep{2018AJ....155..259H,2019MNRAS.484.5608D} have been ongoing in
the microlensing literature for decades without a clear solution and proper
framing of the issue. In this thesis I will revisit these kinds of questions
while taking into account many recent developments in the fields of
computational statistics and machine learning. My main contributions are the
following:
\begin{itemize}
    \item I wrote an open-source Python package \textsf{caustics} which enables fast and
          accurate computation of binary and triple lens microlensing light curves using
          contour integration. The code runs on both CPU and GPU architectures, and
          crucially, for reasons that I will elaborate in subsequent chapters, it
          supports \emph{automatic differentiation} of all outputs with respect to all
          input parameters which enables the use of statistical methods which are orders
          of magnitude more efficient than existing approaches. The code can easily be
          extended to support quadruple lensing and arbitrary intensity profiles for the
          source star.
    \item I have revisited the topic of degeneracies in the posterior probability
          distributions which appear in the context of microlensing and propose solutions
          to these problems.
    \item I have extended the complex polynomial root solver from \citet{Cameron2021}
          such that it can be executed on GPU architectures and that it supports
          automatic differentiation. This enables the evaluation of ~1M lens equation
          solutions for point source binary and triple lenses in seconds using a CPU and
          miliseconds using a GPU.
    \item I propose an approach to statistical inference on a population of microlensing
          events which different from existing methods and has some advantages.
    \item I briefly discuss the issue of correlated noise in microlensing light curves
          and propose ways to account for it using Gaussian Processes.
\end{itemize}

@@

\subsection{Observed flux}
The magnification $A(t)$ is not a direct observable in microlensing events. We
can only measure the flux of the source star which is usually contaminated with
the flux of other nearby stars in the dense stellar field and potentially also
with light from the lens and possible companions to the lens. The observed flux
is thus given by
\begin{equation}
    F(t)=F_S(t)+F_B(t)
    \label{eq:flux_observed}
\end{equation}
where $F_s$ is the flux of the source star and $F_B$ is the contamination flux also
called the \emph{blending flux}. To quantify the strength of the blending e can
define the emph{source flux fraction} $b_S$ as the fraction of total (unmagnified)
flux that is coming from the source star:
\begin{equation}
    b_S\equiv F_S/(F_S + F_B)
\end{equation}
For highly blended events $b_S\approx 0$ and in absence of blending $b_S=1$.
The source flux $F_S$ and the blending flux $F_B$ are generally highly correlated
and it is often preferable to use a slightly different parametrization proposed by
\citet{2009MNRAS.393..816D}. In this parametrization we use the
\emph{baseline flux} $F_\mathrm{base}$ and the difference between
the baseline flux and the flux at peak magnification $\Delta F\equiv F_S[A(u_0)-1]$
as the new parameters.
Equation~\ref{eq:flux_observed} then takes the form
\begin{equation}
    f=\Delta F\frac{A(t) - 1}{A(t_0)-1}+F_\mathrm{base}
    \label{eq:flux_dominik}
\end{equation}
It is generally straightforward to measure the flux difference $\Delta F$, the
baseline flux $F_\mathrm{base}$ and the center of the curve $t_0$ but getting a
good estimate of $t_E$ even outside of the regime $u_0\rightarrow 0$ and
$u_0\gg 1$ requires good photometric sampling along the ``wings'' of the light
curve \citep{2009MNRAS.393..816D}.

@@

Following \citet{2019AJ....157...64L}, we start by setting up a cartesian
coordinate system on the units sphere, such that
\begin{align}
     & x=\sin \theta \cos \phi \\
     & y=\sin \theta \sin \phi \\
     & z=\cos \theta
\end{align}
and the observer is located on the $z$-axis at $\infty$ such that the projected disc
of the body is centered at the origin of the $xy$ plane with the unit vector
$\hat{\mathbf{x}}$ pointing to the right and $\hat{\mathbf{y}}$ pointing up.
We introduce spherical harmonics $Y_{l m}(\theta, \phi)$ of degree $l\geq 0$ and order
$m\in [-l, l]$ defined  as
\begin{equation}
    Y_{l m}(\theta, \phi)= \begin{cases}\bar{P}_{l m}(\cos \theta) \cos (m \phi) & m \geqslant 0 \\ \bar{P}_{l|m|}(\cos \theta) \sin (|m| \phi) & m<0\end{cases}
    \label{eq:spherical_harmonics}
\end{equation}
When the spherical harmonics defined above are rewritten in terms of $x$, $y$, and $z$
they become polynomials in these variables (see Appendix A in \citet{2019AJ....157...64L}).
We can expand the specific intensity distribution $I(x,y)$ in a spherical harmonic basis
as
\begin{equation}
    I(x, y)=\tilde{\mathbf{y}}^{\top}(x, y) \mathbf{y}
    \label{eq:sh_expansion}
\end{equation}
where $\tilde{\mathbf{y}}$ is the basis vector of spherical harmonics  arranged in
increasing order:
\begin{align}
     & \tilde{\mathbf{y}}=\left(\begin{array}{lllll}
                                    Y_{0,0} & Y_{1,-1} & Y_{1,0} & Y_{1,1} & Y_{2,-2}
                                \end{array}\right. \\
     & \left.\begin{array}{lllll}
                 Y_{2,-1} & Y_{2,0} & Y_{2,1} & Y_{2,2} & \cdots
             \end{array}\right)^{\top} \quad,
\end{align}
and $\mathbf{y}$ is a vector of scalar \emph{spherical harmonic coefficients}.
$\mathbf{y}$ is a central quantity which defines the \emph{map} (the specific
intensity at every point) of the body.
\citet{2019AJ....157...64L} shows that we can represent $y$ in a polynomial basis (see
\citet{2019AJ....157...64L} for the general expression)
\begin{equation}
    \tilde{\boldsymbol{p}}=\left(\begin{array}{llllllllll}
        1 & x & z & y & x^{2} & x z & x y & y z & y^{2} & \cdots
    \end{array}\right)^{\top}
\end{equation}
such that
\begin{align}
    I(x, y) & =\tilde{\mathbf{p}}^{\top}(x, y) \mathbf{p}                    \\
            & =\tilde{\mathbf{p}}^{\top}(x, y) \mathbf{A}_1 \mathbf{y}\quad,
    \label{eq:intensity_poly_basis}
\end{align}
and also in the Green's basis
\begin{equation}
    \tilde{\boldsymbol{g}}=\left(\begin{array}{llllllllll}
        1 & 2 x & z & y & 3 x^{2} & -3 x z & 2 x y & 3 y z & y^{2} & \cdots
    \end{array}\right)^{\top}
\end{equation}
such that
\begin{align}
    I(x, y) & =\tilde{\mathbf{g}}^{\top}(x, y) \mathbf{g}              \\
            & =\tilde{\mathbf{g}}^{\top}(x, y) \mathbf{A}_2 \mathbf{p} \\
            & =\tilde{\mathbf{g}}^{\top}(x, y) \mathbf{A} \mathbf{y}
\end{align}
where $\mathbf{A}_1$ is the change-of-basis matrix from $\mathbf{y}$ to $\mathbf{p}$,
$\mathbf{A}_2$ is the change-of-basis matrix from $\mathbf{p}$ to $\mathbf{g}$, and
$\mathbf{A}\equiv \mathbf{A}_1 \mathbf{A}_2$.

To compute rotational light curves we also need to be able to specify the
orientation of a surface map with coefficients $\mathbf y$. We can rotate the
map using the Wigner rotation matrix $\mathbf{R}(\mathrm{I}, \Lambda, \Theta)$
that rotates $y$ given the body's inclination $\mathrm{I}$, obliquity $\Lambda$
and rotational phase $\Theta$. The rotated map is then $\mathbf{R}(\mathrm{I},
    \Lambda, \Theta) \mathbf{y}$ and Equation~\ref{eq:intensity_poly_basis} can be
rewritten more generally as
\begin{equation}
    I(x, y)=\tilde{\mathbf{p}}^{\top}(x, y) \mathbf{A}_1 \mathbf{R}\mathbf{y}
\end{equation}
The total flux measured by an observer is is given by an integral of the specific
intensity over a region S of the projected disc of the body:
\begin{align}
    F & =\oiint I(x, y) \ud S                                                            \\
      & =\oiint \tilde{\mathbf{p}}^{\top}(x, y) \mathbf{A}_{1} \mathbf{R} \mathbf{y} d S \\
      & =\mathbf{r}^{\top} \mathbf{A}_{1} \mathbf{R} \mathbf{y}
\end{align}
where $\mathbf{r}$ is a column vector whose $n$-th element is \citep{2019AJ....157...64L}
\begin{equation}
    r_{n} \equiv \oiint \tilde{p}_{n}(x, y) \ud S
\end{equation}
When the entire disc is visible (no occultor)
\begin{equation}
    r_{n}=\int_{-1}^{1} \int_{-\sqrt{1-x^{2}}}^{\sqrt{1+x^{2}}} \tilde{p}_{n}(x, y) \mathrm{d}y \mathrm{d} x
\end{equation}
and the solution to this integral can be expressed in terms of Gamma functions
\citep[Equation 20 in ][]{2019AJ....157...64L}.
$\mathbf{r}$ and $\mathbf{A}_1$ are independent of the map coefficients $\mathbf{y}$
so they can be pre-computed.

Computing the occultation light curves is more complicated because the the body
is partially occulted by an occultor of radius $r$, centered at $(x_o, y_o)$
and the exposed portion of the disc is a function of $(r, x_o, y_o)$. The
general expression for the flux is
\begin{align}
    F & =\oiint I(x, y) \mathrm{d} S                         \\
      & =\mathbf{s}^\top  \mathbf{A} \mathbf{R} \mathbf{y} .
\end{align}
where the vector
\begin{equation}
    \mathbf{s}^\top \equiv \oiint \tilde{\mathbf{g}}^{\top}(x, y) \mathrm{d} S
\end{equation}
is defined to be the solution to the integral.
\citet{2019AJ....157...64L} solves this integral in two steps.  First, we rotate
the coordinate system about the  $z$-axis  so that the occultor lies at along the
$+y$ axis and its center is at
a distance $b=\sqrt{x_o^{2}+y_o^{2}}$ from the origin. This substantially simplifies
the limits of integration.
The second step is to use Green's theorem to express the surface integral of
$\tilde{\mathbf{g}}_n$ as a line integral of a vector function $\mathbf{G}_n$
along the boundary of the visible portion of the occulted disc.
The $n$-th element of the solution vector $\mathbf{s}^\top$ is
\begin{equation}
    s_{n}=\oiint \tilde{g}_{n}(x, y) \mathrm{d} S=\oint \mathbf{G}_{n}(x, y) \cdot \ud \mathbf{r}
\end{equation}
where
$\mathbf{G}_{n}(x, y)=G_{n x}(x, y) \hat{\mathbf{x}}+G_{n y}(x, y) \hat{\mathbf{y}}$
is constructed such that
\begin{equation}
    \mathbf{D} \wedge \mathbf{G}_{n}=\tilde{g}_{n}(x, y)
    \label{eq:G_n}
\end{equation}
were $\mathbf{D} \wedge \mathbf{G}_{n}$ is the exterior derivative of $\mathbf{G}_n$
given by
\begin{equation}
    \mathbf{D} \wedge \mathbf{G}_{n} \equiv \frac{\ud G_{n y}}{\ud x}-\frac{\ud G_{n_{x}}}{\ud y}
\end{equation}
in Cartesian coordinates.
\citet{2019AJ....157...64L} provides one possible expression for $\mathbf{G}_n$
which satisfies Equation~\ref{eq:G_n} and shows that the final solution can be written as
\begin{equation}
    s_{n}=\mathcal{Q}\left(\mathbf{G}_{n}\right)-\mathcal{P}\left(\mathbf{G}_{n}\right)
\end{equation}
where $\mathcal{P}(\mathbf{G}_n)$ is the  line integral along the arc of the occultor
of radius $r$ and $\mathcal{Q}(\mathbf{G}_n)$ is the line integral along the arc of the
occulted body of unit radius. Solutions to these integrals are long but they consist of
analytic functions such as sines, cosines and complete elliptic integrals and importantly
they only need to be evaluated once for an arbitrary map with fixed geometry of an
occultation \citep{2019AJ....157...64L}.

@@

\subsubsection{Reflected light}
The results derived in the previous section are valid only for light emitted
from the occulted body. Equation~\ref{eq:starry} is a good model for stellar
light curves and secondary eclipse light curves and phase curves of planets
observed in the mid to far infrared wavelengths where the contribution of
scattered starlight to total flux is negligible. Modeling reflected light phase
curves and occultations is far more complicated because one has to take into
account the the nonuniform illumination of the body and the possible presence
of a terminator line (boundary between day and night).
\citet{2021arXiv210306275L} extends the formalism described in the previous
section to the case of reflected light phase curves and occultations for a body
whose spatial \emph{albedo} (specifically the \emph{spherical albedo} -- the
fraction of power incident on a body at a given wavelength that is scattered
into space in all directions) distribution is expanded in a basis of spherical
harmonics. Thus, analogous to Equation~\ref{eq:sh_expansion} we have
\begin{equation}
    A(x, y)=\tilde{\mathbf{y}}^{\top}(x, y) \mathbf{y}
\end{equation}
where $A$ is the albedo. We work under the assumption of isotropic (Lambertian)
scattering of light at every point on the body which means that the illumination
profile $\mathcal{I}$ of the surface is given by Lambert's law:
\begin{equation}
    \mathcal{I}\left(\vartheta_{\mathrm{i}}\right)=\mathcal{I}_{0} \max \left(0, \cos \vartheta_{\mathrm{i}}\right)
\end{equation}
where $\vartheta_i$ is the angle between the incident radiation at the surface normal
and $\mathcal{I}_0$ is the maximum illumination. In case of, for example, a planet illuminated
by its host star \citep[Appendix A.2 in][]{2021arXiv210306275L}
\begin{equation}
    \mathcal{I}_{0}=\frac{f_{s}}{\pi r_{\mathrm{s}}^{2}}
\end{equation}
where $r_s$ is the distance between the planet and the star in units of planet's radius
and $f_s$ is the stellar flux measured at the observer in some arbitrary units.
We assume that $f_s=1$ so that all fluxes are defined as the fraction of the flux
of the illumination source at the observer.
It is the presence of the day/night terminator that complicates the calculation
of total flux the most. The limits of integration end up depending on a
solution to a quartic equation specifying the points of intersection between
the occultor and the day/night terminator line and the solution to those
integrals are a function of \emph{incomplete} elliptic integrals as opposed to
complete elliptic integrals as was the case for occultations in emitted light.
The full derivation of the total flux is provided by \citet{2021arXiv210306275L} and
I will only briefly summarize  the result here for completeness.

@@

the reflected light solution vector $\mathbb{s}^\top$.

@@

    f=\mathbb{s}^\top\left(b, \theta^{\prime}, b_{\mathrm{o}}, r_{\mathrm{o}}\right) \mathbf{A}_{\mathbf{2}} \mathbf{I}\left(b, \theta^{\prime}, r_{\mathrm{s}}\right) \mathbf{A}_{\mathbf{1}} \mathbf{R}^{\prime}\left(x_{\mathrm{o}}, y_{\mathrm{o}}\right) \mathbf{R}(\mathrm{I}, \Lambda, \Theta) \mathbf{y}

@@

    f_{0}=\mathbb{r}^{\top}(b) \mathbf{I}\left(b, \theta^{\prime \prime}, r_{\mathrm{s}}\right) \mathbf{A}_{\mathbf{1}} \mathbf{R}^{\prime \prime}\left(x_{\mathrm{s}}, y_{\mathrm{s}}\right) \mathbf{R}(\mathrm{I}, \Lambda, \Theta) \mathbf{y}

@@

$\mathbb{r}^{\top}(b)$ and $\mathbb{s}^{\top}\left(b, \theta^{\prime}, b_{\mathrm{o}}, r_{\mathrm{o}}\right)$

@@

\subsection{The starry code}
The method described in the previous section is implemented in the
\textsf{Python} package
\textsf{starry}\footnote{\url{https://starry.readthedocs.io/en/latest/}}.
\textsf{starry} is many orders of magnitude faster and more accurate than
similar codes which rely on numerical integration. It enables computation of
rotational light curves, planetary transits, secondary eclipse light curves and
planet-planet occultations in both emitted and reflected light. Recent
extensions to \textsf{starry}, not mentioned in Section~\ref{sec:occultations},
extended the spherical harmonic formalism to the problem of Doppler imaging
\citep{2021arXiv211006271L} and add support for computing occultation light
curves of oblate bodies \citep{2022ApJ...925..185D}. In
Chapters~\ref{ch:mapping_io} and \ref{ch:mapping_exoplanets} we make extensive
use of \textsf{starry} to explore solutions to the \emph{inverse problem} --
how can we obtain an estimate of the map coefficients $\mathbf{y}$ which best
explain the observed light curve? \textsf{starry} abstracts the complexities of
computing the total flux and allows us to treat this problem as a linear
problem of the form
\begin{equation}
    f = \mathbf{X}\mathbf{y}
\end{equation}
where  $\mathbf{X}$ is computed by \textsf{starry} depending on the geometry of an
occultation or transit event. The inverse problem is highly degenerate
and making progress requires  imposing some prior structure on the solution
$\mathbf{y}$.

@@

\end{document}ii

>>>>

% Define document class
\documentclass[12pt,dvipsnames]{report}
\usepackage{aas_macros}

@@

% Figures & tables 
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs} % much better looking tables

@@

% Algorithms 
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

% Links 
\hypersetup{
    hidelinks,
    colorlinks=true,
    allcolors=NavyBlue,
}

@@

% Custom command
\newcommand{\ud}{\,\mathrm{d}}
\renewcommand{\vec}[1]{\boldsymbol{\mathbf{#1}}}
\newcommand{\hquad}{~~}

@@

The assumptions that existing methods for modeling microlensing events rely on
are often opaque and unquestioned. Discussions on model ``degeneracies''
\citep{2014MNRAS.437.4006S,2019AJ....157...23H,2018AcA....68...43S,2009MNRAS.393..816D},
correlated noise \citep{2015ApJ...812..136B,2019MNRAS.488.3308L} and model
comparison \citep{2018AJ....155..259H,2019MNRAS.484.5608D} have been ongoing in
the microlensing literature for decades without a clear solution and proper
framing of the issue. In this thesis I will revisit these kinds of questions
while taking into account many recent developments in the fields of
computational statistics and machine learning. My main contributions are the
following:
\begin{itemize}
    \item I wrote an open-source Python package \textsf{caustics} which enables fast and
          accurate computation of binary and triple lens microlensing light curves using
          contour integration. The code runs on both CPU and GPU architectures, and
          crucially, for reasons that I will elaborate in subsequent chapters, it
          supports \emph{automatic differentiation} of all outputs with respect to all
          input parameters which enables the use of statistical methods which are orders
          of magnitude more efficient than existing approaches. The code can easily be
          extended to support quadruple lensing and arbitrary intensity profiles for the
          source star.
    \item I have revisited the topic of degeneracies in the posterior probability
          distributions which appear in the context of microlensing and propose solutions
          to these problems.
    \item I have extended the complex polynomial root solver from \citet{Cameron2021}
          such that it can be executed on GPU architectures and that it supports
          automatic differentiation. This enables the evaluation of ~1M lens equation
          solutions for point source binary and triple lenses in seconds using a CPU and
          miliseconds using a GPU.
    \item I propose an approach to statistical inference on a population of microlensing
          events which different from existing methods and has some advantages.
    \item I briefly discuss the issue of correlated noise in microlensing light curves
          and propose ways to account for it using Gaussian processes.
\end{itemize}

@@

\subsection{Observed flux}
The magnification $A(t)$ is not a direct observable in microlensing events. We
can only measure the flux of the source star which is usually contaminated with
the flux of other nearby stars in the dense stellar field and potentially also
with light from the lens and possible companions to the lens. The observed flux
is thus given by
\begin{equation}
    F(t)=F_S\,A(t)+F_B
    \label{eq:flux_observed_micro}
\end{equation}
where $F_s$ is the flux of the source star and $F_B$ is the contamination flux also
called the \emph{blending flux}. To quantify the strength of the blending e can
define the emph{source flux fraction} $b_S$ as the fraction of total (unmagnified)
flux that is coming from the source star:
\begin{equation}
    b_S\equiv F_S/(F_S + F_B)
\end{equation}
For highly blended events $b_S\approx 0$ and in absence of blending $b_S=1$.
The source flux $F_S$ and the blending flux $F_B$ are generally highly correlated
and it is often preferable to use a slightly different parametrization proposed by
\citet{2009MNRAS.393..816D}. In this parametrization we use the
\emph{baseline flux} $F_\mathrm{base}$ and the difference between
the baseline flux and the flux at peak magnification $\Delta F\equiv F_S[A(u_0)-1]$
as the new parameters.
Equation~\ref{eq:flux_observed_micro} then takes the form
\begin{equation}
    f=\Delta F\frac{A(t) - 1}{A(t_0)-1}+F_\mathrm{base}
    \label{eq:flux_dominik}
\end{equation}
It is generally straightforward to measure the flux difference $\Delta F$, the
baseline flux $F_\mathrm{base}$ and the center of the curve $t_0$ but getting a
good estimate of $t_E$ even outside of the regime $u_0\rightarrow 0$ and
$u_0\gg 1$ requires good photometric sampling along the ``wings'' of the light
curve \citep{2009MNRAS.393..816D}.

@@

Following \citet{2019AJ....157...64L}, we start by setting up a cartesian
coordinate system on the units sphere, such that
\begin{align}
     & x=\sin \theta \cos \phi \\
     & y=\sin \theta \sin \phi \\
     & z=\cos \theta
\end{align}
and the observer is located on the $z$-axis at $\infty$ such that the projected disc
of the body is centered at the origin of the $xy$ plane with the unit vector
$\hat{\mathbf{x}}$ pointing to the right and $\hat{\mathbf{y}}$ pointing up.
We introduce spherical harmonics $Y_{l m}(\theta, \phi)$ of degree $l\geq 0$ and order
$m\in [-l, l]$ defined  as
\begin{equation}
    Y_{l m}(\theta, \phi)= \begin{cases}\bar{P}_{l m}(\cos \theta) \cos (m \phi) & m \geqslant 0 \\ \bar{P}_{l|m|}(\cos \theta) \sin (|m| \phi) & m<0\end{cases}
    \label{eq:spherical_harmonics}
\end{equation}
When the spherical harmonics defined above are rewritten in terms of $x$, $y$, and $z$
they become polynomials in these variables (see Appendix A in \citet{2019AJ....157...64L}).
We can expand the specific intensity distribution $I(x,y)$ in a spherical harmonic basis
as
\begin{equation}
    I(x, y)=\tilde{\mathbf{y}}^{\intercal}(x, y) \mathbf{y}
    \label{eq:sh_expansion}
\end{equation}
where $\tilde{\mathbf{y}}$ is the basis vector of spherical harmonics  arranged in
increasing order:
\begin{align}
     & \tilde{\mathbf{y}}=\left(\begin{array}{lllll}
                                    Y_{0,0} & Y_{1,-1} & Y_{1,0} & Y_{1,1} & Y_{2,-2}
                                \end{array}\right. \\
     & \left.\begin{array}{lllll}
                 Y_{2,-1} & Y_{2,0} & Y_{2,1} & Y_{2,2} & \cdots
             \end{array}\right)^{\intercal} \quad,
\end{align}
and $\mathbf{y}$ is a vector of scalar \emph{spherical harmonic coefficients}.
$\mathbf{y}$ is a central quantity which defines the \emph{map} (the specific
intensity at every point) of the body.
\citet{2019AJ....157...64L} shows that we can represent $y$ in a polynomial basis (see
\citet{2019AJ....157...64L} for the general expression)
\begin{equation}
    \tilde{\boldsymbol{p}}=\left(\begin{array}{llllllllll}
        1 & x & z & y & x^{2} & x z & x y & y z & y^{2} & \cdots
    \end{array}\right)^{\intercal}
\end{equation}
such that
\begin{align}
    I(x, y) & =\tilde{\mathbf{p}}^{\intercal}(x, y) \mathbf{p}                    \\
            & =\tilde{\mathbf{p}}^{\intercal}(x, y) \mathbf{A}_1 \mathbf{y}\quad,
    \label{eq:intensity_poly_basis}
\end{align}
and also in the Green's basis
\begin{equation}
    \tilde{\boldsymbol{g}}=\left(\begin{array}{llllllllll}
        1 & 2 x & z & y & 3 x^{2} & -3 x z & 2 x y & 3 y z & y^{2} & \cdots
    \end{array}\right)^{\intercal}
\end{equation}
such that
\begin{align}
    I(x, y) & =\tilde{\mathbf{g}}^{\intercal}(x, y) \mathbf{g}              \\
            & =\tilde{\mathbf{g}}^{\intercal}(x, y) \mathbf{A}_2 \mathbf{p} \\
            & =\tilde{\mathbf{g}}^{\intercal}(x, y) \mathbf{A} \mathbf{y}
\end{align}
where $\mathbf{A}_1$ is the change-of-basis matrix from $\mathbf{y}$ to $\mathbf{p}$,
$\mathbf{A}_2$ is the change-of-basis matrix from $\mathbf{p}$ to $\mathbf{g}$, and
$\mathbf{A}\equiv \mathbf{A}_1 \mathbf{A}_2$.

To compute rotational light curves we also need to be able to specify the
orientation of a surface map with coefficients $\mathbf y$. We can rotate the
map using the Wigner rotation matrix $\mathbf{R}(\mathrm{I}, \Lambda, \Theta)$
that rotates $y$ given the body's inclination $\mathrm{I}$, obliquity $\Lambda$
and rotational phase $\Theta$. The rotated map is then $\mathbf{R}(\mathrm{I},
    \Lambda, \Theta) \mathbf{y}$ and Equation~\ref{eq:intensity_poly_basis} can be
rewritten more generally as
\begin{equation}
    I(x, y)=\tilde{\mathbf{p}}^{\intercal}(x, y) \mathbf{A}_1 \mathbf{R}\mathbf{y}
\end{equation}
The total flux measured by an observer is is given by an integral of the specific
intensity over a region S of the projected disc of the body:
\begin{align}
    F & =\oiint I(x, y) \ud S                                                                 \\
      & =\oiint \tilde{\mathbf{p}}^{\intercal}(x, y) \mathbf{A}_{1} \mathbf{R} \mathbf{y} d S \\
      & =\mathbf{r}^{\intercal} \mathbf{A}_{1} \mathbf{R} \mathbf{y}
\end{align}
where $\mathbf{r}$ is a column vector whose $n$-th element is \citep{2019AJ....157...64L}
\begin{equation}
    r_{n} \equiv \oiint \tilde{p}_{n}(x, y) \ud S
\end{equation}
When the entire disc is visible (no occultor)
\begin{equation}
    r_{n}=\int_{-1}^{1} \int_{-\sqrt{1-x^{2}}}^{\sqrt{1+x^{2}}} \tilde{p}_{n}(x, y) \mathrm{d}y \mathrm{d} x
\end{equation}
and the solution to this integral can be expressed in terms of Gamma functions
\citep[Equation 20 in ][]{2019AJ....157...64L}.
$\mathbf{r}$ and $\mathbf{A}_1$ are independent of the map coefficients $\mathbf{y}$
so they can be pre-computed.

Computing the occultation light curves is more complicated because the the body
is partially occulted by an occultor of radius $r$, centered at $(x_o, y_o)$
and the exposed portion of the disc is a function of $(r, x_o, y_o)$. The
general expression for the flux is
\begin{align}
    F & =\oiint I(x, y) \mathrm{d} S                              \\
      & =\mathbf{s}^\intercal  \mathbf{A} \mathbf{R} \mathbf{y} .
\end{align}
where the vector
\begin{equation}
    \mathbf{s}^\intercal \equiv \oiint \tilde{\mathbf{g}}^{\intercal}(x, y) \mathrm{d} S
\end{equation}
is defined to be the solution to the integral.
\citet{2019AJ....157...64L} solves this integral in two steps.  First, we rotate
the coordinate system about the  $z$-axis  so that the occultor lies at along the
$+y$ axis and its center is at
a distance $b=\sqrt{x_o^{2}+y_o^{2}}$ from the origin. This substantially simplifies
the limits of integration.
The second step is to use Green's theorem to express the surface integral of
$\tilde{\mathbf{g}}_n$ as a line integral of a vector function $\mathbf{G}_n$
along the boundary of the visible portion of the occulted disc.
The $n$-th element of the solution vector $\mathbf{s}^\intercal$ is
\begin{equation}
    s_{n}=\oiint \tilde{g}_{n}(x, y) \mathrm{d} S=\oint \mathbf{G}_{n}(x, y) \cdot \ud \mathbf{r}
\end{equation}
where
$\mathbf{G}_{n}(x, y)=G_{n x}(x, y) \hat{\mathbf{x}}+G_{n y}(x, y) \hat{\mathbf{y}}$
is constructed such that
\begin{equation}
    \mathbf{D} \wedge \mathbf{G}_{n}=\tilde{g}_{n}(x, y)
    \label{eq:G_n}
\end{equation}
were $\mathbf{D} \wedge \mathbf{G}_{n}$ is the exterior derivative of $\mathbf{G}_n$
given by
\begin{equation}
    \mathbf{D} \wedge \mathbf{G}_{n} \equiv \frac{\ud G_{n y}}{\ud x}-\frac{\ud G_{n_{x}}}{\ud y}
\end{equation}
in Cartesian coordinates.
\citet{2019AJ....157...64L} provides one possible expression for $\mathbf{G}_n$
which satisfies Equation~\ref{eq:G_n} and shows that the final solution can be written as
\begin{equation}
    s_{n}=\mathcal{Q}\left(\mathbf{G}_{n}\right)-\mathcal{P}\left(\mathbf{G}_{n}\right)
\end{equation}
where $\mathcal{P}(\mathbf{G}_n)$ is the  line integral along the arc of the occultor
of radius $r$ and $\mathcal{Q}(\mathbf{G}_n)$ is the line integral along the arc of the
occulted body of unit radius. Solutions to these integrals are long but they consist of
analytic functions such as sines, cosines and complete elliptic integrals and importantly
they only need to be evaluated once for an arbitrary map with fixed geometry of an
occultation \citep{2019AJ....157...64L}.

@@

\subsubsection{Reflected light}
The results derived in the previous section are valid only for light emitted
from the occulted body. Equation~\ref{eq:starry} is a good model for stellar
light curves and secondary eclipse light curves and phase curves of planets
observed in the mid to far infrared wavelengths where the contribution of
scattered starlight to total flux is negligible. Modeling reflected light phase
curves and occultations is far more complicated because one has to take into
account the the nonuniform illumination of the body and the possible presence
of a terminator line (boundary between day and night).
\citet{2021arXiv210306275L} extends the formalism described in the previous
section to the case of reflected light phase curves and occultations for a body
whose spatial \emph{albedo} (specifically the \emph{spherical albedo} -- the
fraction of power incident on a body at a given wavelength that is scattered
into space in all directions) distribution is expanded in a basis of spherical
harmonics. Thus, analogous to Equation~\ref{eq:sh_expansion} we have
\begin{equation}
    A(x, y)=\tilde{\mathbf{y}}^{\intercal}(x, y) \mathbf{y}
\end{equation}
where $A$ is the albedo. We work under the assumption of isotropic (Lambertian)
scattering of light at every point on the body which means that the illumination
profile $\mathcal{I}$ of the surface is given by Lambert's law:
\begin{equation}
    \mathcal{I}\left(\vartheta_{\mathrm{i}}\right)=\mathcal{I}_{0} \max \left(0, \cos \vartheta_{\mathrm{i}}\right)
\end{equation}
where $\vartheta_i$ is the angle between the incident radiation at the surface normal
and $\mathcal{I}_0$ is the maximum illumination. In case of, for example, a planet illuminated
by its host star \citep[Appendix A.2 in][]{2021arXiv210306275L}
\begin{equation}
    \mathcal{I}_{0}=\frac{f_{s}}{\pi r_{\mathrm{s}}^{2}}
\end{equation}
where $r_s$ is the distance between the planet and the star in units of planet's radius
and $f_s$ is the stellar flux measured at the observer in some arbitrary units.
We assume that $f_s=1$ so that all fluxes are defined as the fraction of the flux
of the illumination source at the observer.
It is the presence of the day/night terminator that complicates the calculation
of total flux the most. The limits of integration end up depending on a
solution to a quartic equation specifying the points of intersection between
the occultor and the day/night terminator line and the solution to those
integrals are a function of \emph{incomplete} elliptic integrals as opposed to
complete elliptic integrals as was the case for occultations in emitted light.
The full derivation of the total flux is provided by \citet{2021arXiv210306275L} and
I will only briefly summarize  the result here for completeness.

@@

the reflected light solution vector $\mathbb{s}^\intercal$.

@@

    f=\mathbb{s}^\intercal\left(b, \theta^{\prime}, b_{\mathrm{o}}, r_{\mathrm{o}}\right) \mathbf{A}_{\mathbf{2}} \mathbf{I}\left(b, \theta^{\prime}, r_{\mathrm{s}}\right) \mathbf{A}_{\mathbf{1}} \mathbf{R}^{\prime}\left(x_{\mathrm{o}}, y_{\mathrm{o}}\right) \mathbf{R}(\mathrm{I}, \Lambda, \Theta) \mathbf{y}

@@

    f_{0}=\mathbb{r}^{\intercal}(b) \mathbf{I}\left(b, \theta^{\prime \prime}, r_{\mathrm{s}}\right) \mathbf{A}_{\mathbf{1}} \mathbf{R}^{\prime \prime}\left(x_{\mathrm{s}}, y_{\mathrm{s}}\right) \mathbf{R}(\mathrm{I}, \Lambda, \Theta) \mathbf{y}

@@

$\mathbb{r}^{\intercal}(b)$ and $\mathbb{s}^{\intercal}\left(b, \theta^{\prime}, b_{\mathrm{o}}, r_{\mathrm{o}}\right)$

@@

\subsection{The starry code}
The method described in the previous section is implemented in the
\textsf{Python} package
\textsf{starry}\footnote{\url{https://starry.readthedocs.io/en/latest/}}.
\textsf{starry} is many orders of magnitude faster and more accurate than
similar codes which rely on numerical integration. It enables computation of
rotational light curves, planetary transits, secondary eclipse light curves and
planet-planet occultations in both emitted and reflected light. Recent
extensions to \textsf{starry}, not mentioned in Section~\ref{sec:occultations},
extended the spherical harmonic formalism to the problem of Doppler imaging
\citep{2021arXiv211006271L} and add support for computing occultation light
curves of oblate bodies \citep{2022ApJ...925..185D}. In
Chapters~\ref{ch:mapping_io} and \ref{ch:mapping_exoplanets} we make extensive
use of \textsf{starry} to explore solutions to the \emph{inverse problem} --
how can we obtain an estimate of the map coefficients $\mathbf{y}$ which best
explain the observed light curve? \textsf{starry} abstracts the complexities of
computing the total flux and allows us to treat this problem as a linear
problem of the form
\begin{equation}
    f = \mathbf{X}\mathbf{y}
    \label{eq:starry_linear_model}
\end{equation}
where  $\mathbf{X}$ is computed by \textsf{starry} depending on the geometry of an
occultation or transit event. The inverse problem is highly degenerate
and making progress requires  imposing some prior structure on the solution
$\mathbf{y}$.

@@

\end{document}
